---
title: "R Notebook"
output: html_notebook
---
*Run* : *Cmd+Shift+Enter*
*Insert Chunk*: *Cmd+Option+I*
*Preview*: *Cmd+Shift+K* 

```{r}
setwd("~/savio/global/scratch/stacy-l/spermSV/notebooks")
library(tidyverse)
library(paletteer)
library(svglite)
library(viridis)
library(vcfR)
library(plyr)
```

Use repeatmasker `.out` file to obtain a mapping table of motif to class:

```{r}
repeat_out <- read_table("../output_t2t/sniffles/mosaic/multi_sample/repeatmasker/multi_sample.filtered.fa.out", skip = 3, col_names = FALSE)
motif_class = repeat_out %>%
  select(X10, X11) %>%
  rename('motif' = X10, 'class' = X11) %>%
  distinct() # rename, drop all duplicate entries
```

Load in `.gff` file and clean.

```{r}
repeat_gff <- read_table("../output_t2t/sniffles/mosaic/multi_sample/repeatmasker/multi_sample.filtered.fa.out.gff", skip = 3, col_names = FALSE)
colnames(repeat_gff) <- c('ID', 'source', 'feature', 'start', 'end', 'score', 'strand', 'frame', 'attribute', 'motif', 'unknown1', 'unknown2') # from gff format, padded with motif, unknown1, unknown2 (because I don't know what those are)
repeat_gff <- repeat_gff %>% 
  select(-c('attribute', 'unknown1', 'unknown2')) # drop out of frame
repeat_gff$motif <- repeat_gff$motif %>% 
  str_extract('(?<=:)(.*)(?=\")') # regex out motif name

```

Merge `.gff` table and motif-class mapping table.

```{r}
gff_anno <- merge(repeat_gff, motif_class, by = 'motif')
gff_anno <- gff_anno %>%
  mutate(length = end-start)
```

Create "simplified" labels for classes.

```{r}
gff_anno$broad_class = str_replace(gff_anno$class,'(?=/)(.*)', '')
```

Raw count histogram:
```{r}
raw_count_hist <- ggplot(gff_anno, aes(fill = broad_class, x = length)) + 
  geom_histogram(color = "#2b2b2b") +
  xlab("Annotated length") +
  ylab("Count") +
  labs(title = "Repeat classes by annotated length",
  subtitle = "(with raw binned counts)",
  fill = "Broad repeat classes") +
  scale_fill_manual(values = paletteer_d("ghibli::PonyoMedium", type = 'continuous', n = 8, direction = -1)) +
  theme_classic() +
  theme(text=element_text(family="Helvetica Neue"),
        plot.title = element_text(size=18))

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_hist_raw.svg",
  plot = raw_count_hist,
  device = 'svg',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_hist_raw.png",
  plot = raw_count_hist,
  device = 'png',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)

# no PDF for now because of font issue
```

With sqrt count (y axis) transformation:
```{r}
sqrt_count_hist <- ggplot(gff_anno, aes(fill = broad_class, x = length)) + 
  geom_histogram(color = "#2b2b2b") +
  xlab("Annotated length") +
  labs(fill = "Broad repeat classes") +
  scale_fill_manual(values = paletteer_d("ghibli::PonyoMedium", type = 'continuous', n = 8, direction = -1)) +
  scale_y_sqrt() +
  ylab("*Count (sqrt transform)") +
  labs(title = "Repeat classes by annotated length",
  subtitle = "(with sqrt-transformed counts)",
  fill = "Broad repeat classes") +
  theme_classic() +
  theme(text=element_text(family="Helvetica Neue"),
        plot.title = element_text(size=18))

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_hist_sqrt.svg",
  plot = sqrt_count_hist,
  device = 'svg',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_hist_sqrt.png",
  plot = sqrt_count_hist,
  device = 'png',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)
```

Plotting counts of classes overall:
```{r}
class_count <- gff_anno %>%
  group_by(class) %>%
  summarise(count = n())
class_count <- class_count[order(-class_count$count),]

class_count_plot <-ggplot(class_count, aes(fill = class, x = reorder(class, -count), y = count)) + 
  geom_bar(color = "#2b2b2b", stat = 'identity', show.legend = FALSE) +
  xlab("Repeat class") +
  ylab("Count") +
  labs(title = "Repeat classes by count",
  subtitle = "Denoted with specific repeat classes",
  fill = "Specific repeat classes") +
  theme_classic() +
  theme(text=element_text(family="Helvetica Neue"),
        plot.title = element_text(size=18),
        axis.text.x = element_text(angle=45, hjust = 1)) +
  scale_fill_manual(values = paletteer_d("LaCroixColoR::Pure", type = 'continuous', n = 20, direction = 1))

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_count.svg",
  plot = class_count_plot,
  device = 'svg',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)

ggsave(
  filename="sniffles2_mosaic_repeatmasker_alt_count.png",
  plot = class_count_plot,
  device = 'png',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 6,
  units = "in",
  dpi = 300
)

# no PDF for now because of font issue
  
```

Ugh, more plots to jazz up the presentation. Look at the actual VCF...

```{r}
filtered <- read.vcfR("output_t2t/sniffles/mosaic/multi_sample/ucsc_variant_annotation/multi_sample.filtered.with.header.vcf", verbose = FALSE)
```

```{r}
gt_raw <- extract.gt(filtered, element="GT") # in phased format
gt <- extract.gt(filtered, element="GT", as.numeric=TRUE)
rownames(gt) <- 1:nrow(gt)

# heatmap of calls
vcf_heatmap <- heatmap.bp(gt)

ggsave(
  filename="sniffles2_mosaic_vcf_heatmap.svg",
  plot = vcf_heatmap,
  device = 'svg',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 4,
  units = "in",
  dpi = 300
)


ggsave(
  filename="sniffles2_mosaic_vcf_heatmap.png",
  plot = vcf_heatmap,
  device = 'png',
  path = "/Users/stacyli/Documents/Lab Meetings/20230809_assets/",
  scale = 1,
  width = 8,
  height = 4,
  units = "in",
  dpi = 300
)
```


Now, on to the variant predicted effect table from UCSC...

```{r}
# can't think of a good way to deal with this rn so skipping it

ucsc_anno <- read.table("../output_t2t/sniffles/mosaic/multi_sample/ucsc_variant_annotation/20230808_sniffles_mosaic_vep_chr1_to_6.tsv", sep = '\t', skip = 5, header = TRUE)
```

